Intrusion Detection with \\ Genetic Algorithms and Fuzzy Logic

### 2013-12-07 Talk/Presentation.tex
Made a few changes.

  	\begin{itemize}
  		\item Trying to determine version of software is being run, if it has known issue it allows attacker to attempt to attack that.
  	\end{itemize}
  \end{itemize}
\end{frame}

@@

\subsection{Determining the Accuracy of an Algorithm}
\begin{frame}
  \frametitle{Determining the Accuracy of an Algorithm}
\begin{table}
Predicted
\begin{tabular}{l|ll}
Actual   & Not Attack & Attack \\ \hline
Not Attack & True Negative (TN) & False Positive (FP) \\
Attack & False Negative (FN) & True Positive (TP)  \\
\end{tabular}
\end{table}
\linespace
\linespace
\begin{center}
Detection rate (DR): percentage of normal and attack activity correctly classified from the total number of data records.
\end{center}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Fuzzy Classification]{Fuzzy Classification}
\subsection{Fuzzy Logic}
\begin{frame}
	\frametitle{Fuzzy Logic}
	\begin{itemize}
		\item Fuzzy logic: used in intrusion detection systems to find degree of certainty of record being attack.
			\begin{itemize}
				\item If it's not clear if activity is an attack, fuzzy logic says where on a spectrum it is and how certain it is of being an attack.
			\end{itemize}
		\linespace
		\linespace
		\linespace
		\item Fuzzy logic rules: similar to rules described before, except that
consequence is certainty factor. 
		\begin{itemize}
			\item If (\emph{duration} = 6.2) then (\emph{the degree of certainty of the record being an attack is 0.8}).
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\subsection{Encoding of Features and Rules}
\begin{frame}
	\frametitle{Encoding of Features and Rules}
	\begin{itemize}
	\item Four parameters are encoded into blocks. Each block is feature with values 0-7.
	\item Rule has 1 block for each of 12 features followed at end by marker indicating type of attack.
	\end{itemize}

@@

Degree of certainty is computed for each of 12 blocks, if sum of those is greater than a threshold, declared as attack.
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Genetic Algorithms]{Genetic Algorithms}
\subsection{GA Overview}
\begin{frame}
  \frametitle{Genetic Algorithms}
	\begin{itemize}
		\item GAs: search technique used to find solutions to problems.
        \item Possible solutions to problems: represented in a variety of problem dependent ways.

@@

        \linespace
        \item A randomly generated population of potential solutions is created. Mutation, crossover, selection applied to each generation until acceptable solution is found or time limit is exceeded.
	\end{itemize}
\end{frame}

@@

\linespace
Testing has:
\begin{itemize}
	\item Higher DR: IDS is more likely to classify attacks because it can match either attack type.
	\item Lower FN: IDS is less likely to predict that it's not an attack when it really is.
\end{itemize}
\end{frame}

>>>>

  	\begin{itemize}
  		\item Trying to determine what version of software is being run, if that version has known issue, it allows attacker to attempt to attack that.
  	\end{itemize}
  \end{itemize}
\end{frame}

@@

\subsection{Determining the Accuracy of an Algorithm}
\begin{frame}
  \frametitle{Determining the Accuracy of an Algorithm}
\begin{table}
Predicted
\begin{tabular}{l|ll}
Actual   & Not Attack & Attack \\ \hline
Not Attack & True Negative (TN) & False Positive (FP) \\
Attack & False Negative (FN) & True Positive (TP)  \\
\end{tabular}
\end{table}
\linespace
\linespace
\begin{center}
Detection rate (DR): percentage of normal and attack activity correctly classified from the total number of data records.
\end{center}
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Fuzzy Classification]{Fuzzy Classification}
\subsection{Fuzzy Logic}
\begin{frame}
	\frametitle{Fuzzy Logic}
	\begin{itemize}
		\item Fuzzy logic: used in intrusion detection systems to find degree of certainty of a record being attack.
			\begin{itemize}
				\item If it's not clear if the activity is an attack, fuzzy logic says where on a spectrum it is and how certain it is of being an attack.
			\end{itemize}
		\linespace
		\linespace
		\linespace
		\item Fuzzy logic rules: similar to rules described before, except that
consequence is certainty factor. 
		\begin{itemize}
			\item If (\emph{duration} = 6.2) then (\emph{the degree of certainty of the record being an attack is 0.8}).
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\subsection{Encoding of Features and Rules}
\begin{frame}
	\frametitle{Encoding of Features and Rules}
	\begin{itemize}
	\item Four parameters are encoded into blocks. Each block is a feature with values 0-7.
	\item A rule has 1 block for each of 12 features followed at end by a marker indicating type of attack.
	\end{itemize}

@@

Degree of certainty is computed for each of the 12 blocks, if sum of those is greater than a threshold, declared as an attack.
\end{frame}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section[Genetic Algorithms]{Genetic Algorithms}
\subsection{GA Overview}
\begin{frame}
  \frametitle{Genetic Algorithms}
	\begin{itemize}
		\item GAs: search technique used to find solutions to problems.
        \item Possible solutions to problems: represented in a variety of problem dependent ways.

@@

        \linespace
        \item A randomly generated population of potential solutions is created. Mutation, crossover, selection are applied to each generation until acceptable solution is found or time limit is exceeded.
	\end{itemize}
\end{frame}

@@

\linespace
Testing has:
\begin{itemize}
	\item Both DoS and Probe. If it sees either type of attack, it just classifies it as an attack.
	\item Higher DR: IDS is more likely to classify attacks because it can match either attack type.
	\item Lower FN: IDS is less likely to predict that it's not an attack when it really is.
\end{itemize}
\end{frame}
